---
title: "ESM 206 Assignment 4 Report"
author: "Qusai Bhaijeewala and Ruiheng Jin"
date: "11/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

```{r}

# attaching packages of tidyverse, janitor, and DT for data tables.
library(tidyverse)
library(janitor)
library(kableExtra)
library(effsize)


# read in lobster abundance data
# change all the -99999 in the cells to na
# turn all the headers to snake case
lob_abun <- read_csv("lobster_abundance_sbc_lter.csv",
                     na = "-99999") %>% 
  clean_names()
```
## INTRODUCTION

In 2012, the Santa Barbara Coastal Long Term Ecological Research program (SBC LTER) started collecting data on spiny lobster to determine if their populations and sizes were effected by fishing intensity. The purpose of this report is to determine the changes in lobster abudance and lobster size from 2012 to 2018 from five sites in Santa Barbara County. The five sites are Arroyo Quemado, Carpinteria, Isla Vista, Mohawk, Naples. Special attention will be given to the two Marine Protected Area (MPA) of Naples and Isla Vista. Additionally, this report will compare average lobster size in milimeters for MPA and NON MPA sites in the years 2012 and 2018.These finding will pave the way for further analysis that determines the cause of certain trends in lobster population and size changes.

## DATA AND METHODS
The data used in this report was collected by researchers from the Santa Barbara Coastal Long Term Ecological Research program (SBC LTER). The data includes observations of California spiny lobsters at five locations, two of which are Marine Protected Area (MPA). Three variables are studied in this report: year of observation, site of observation, size of lobster (mm). Two-sample, two-sided, unpaired tests (α = 0.05) and cohen’s d are performed for comparisons of mean lobster sizes between MPA sites and non-MPA sites, and between 2012 and 2018. Data analysis is performed using RStudio (Version 1.2.1335).

## RESULTS {.tabset .tabset-fade}

### RESULTS A
#### **CHANGES IN ANNUAL LOBSTER ABUNDANCE AT FIVE SITES**

Isla Vista has the most noticable increase in lobster abundance from 2012 to 2018. Additionally, further analysis is needed to explain the sudden spike in lobster abundance in Isla Vista after 2016. Naples and Mohawk have similar increasing trends, although Naples experienced an increase between 2017 and 2018, while Mohawk a decrease. Carpinteria experienced a large increase between 2016 and 2017, but also a large decrease the following year. The  The Arroyo Quemado's abundance remained low throughout the 2012-2018 period (Figure 1).

```{r}
# data structure: each site has several transects, each transect has ABCD replicates

# uncount lobsters so each lobster has its own row
# Create a new row that says TRUE or FALSE for MPA Sites which are IVEE and NAPL
lob_tidy <- lob_abun %>% 
  tidyr::uncount(count) %>% 
  mutate(mpa = site %in% c("IVEE", "NAPL"))

# making a new column changing TRUE or FALSE to MPA and non-MPA
lob_tidy$mpa_status <- factor(lob_tidy$mpa, labels = c("non-MPA", "MPA"))

# making a new column that gives the full names of the sites so that we can use this column later in graphs

lob_tidy$site_name <- factor(lob_tidy$site, labels = c("Arroyo Quemado", "Carpinteria", "Isla Vista", "Mohawk", "Naples"))

```

```{r}
# count total lobster count by year and site
lob_yr_site <- lob_tidy %>%
  count(year, site) %>% 
  mutate(mpa = site %in% c("IVEE", "NAPL"),
         mpa_status = factor(mpa, labels = c("non-MPA", "MPA")),
         site_name = factor(site, labels = c("Arroyo Quemado", "Carpinteria", "Isla Vista", "Mohawk", "Naples")))  

ggplot(lob_yr_site, aes(x = year, y = n)) +
  facet_wrap(site_name~mpa_status) +
  geom_line(color = "#b62b2b", size = 1) +
  geom_point(color = "#4b1313", size = 2) +
  labs(x = "", 
       y = "Lobster abundance", 
       title = "Lobster Abundance at Five Sites (2012-2018)",
       caption = "Figure 1. Lobster abundance from 5 study sites, from year 2012 to 2018.") +
  scale_x_continuous(expand = expand_scale(add = c(1, 1)),
                     breaks = c(2012, 2014, 2016, 2018)) +
  scale_y_continuous(expand = expand_scale(add = c(0, 0)),
                     limits = c(0, 1000)) +
  theme_light() +
  theme(plot.caption = element_text(hjust = 0, size = 10),
        title = element_text(size = 11)) 
```


### RESULTS B
#### **SHIFTS IN LOBSTER SIZE DISTRIBUTION FROM 2012 TO 2018**

There are two marine protected area (MPA), Naples and Isla Vista. Naples has a higher percentage of larger lobsters in 2018 than in 2012. Ilsa Vista, another MPA site, has an even bigger shift is lobster size. There are a lot of smaller lobsters in 2012, but they become much bigger in 2018. The change in size distributions for Naples and Ilsa Vista makes sense because these sites are marine protected areas, these lobsters have been allowed to grow old and larger in size. For the other three sites, there is a shift towards larger lobster over time, but the shift is much smaller. The biggest shift in lobster size over time for the non mpa sites is seen at the site Carpinteria (Figure 2).

```{r}
# data structure: each site has several transects, and within each transect they have either A, B, C, or D replicates. The count column can has multiple counts of the same transect number and replicate, we want to make each of those counts a seperate row.

# uncount lobsters so each lobster has its own row
lob_tidy2 <- lob_abun %>% 
  tidyr::uncount(count) %>% 
  mutate(mpa = site %in% c("IVEE", "NAPL"))

# making a new column changing TRUE or FALSE to MPA and non-MPA
lob_tidy2$mpa_status <- factor(lob_tidy2$mpa, labels = c("non-MPA", "MPA"))
```


```{r}

# Making a new column that gives the full names of the sites so that we can use this column later in graphs

lob_tidy2$site_name <- factor(lob_tidy2$site, labels = c("Arroyo Quemado", "Carpinteria", "Isla Vista", "Mohawk", "Naples"))

# filtering years of 2012 and 2018 only
lob_size <- lob_tidy2 %>% 
  filter(year %in% c(2012,2018))

#graphing size distributions of lobster size at five sites from two years, 2012, 2018

ggplot(lob_size, aes(x = size_mm)) +
  geom_density(aes(fill = site_name)) +
  facet_grid(site ~ year) +
  theme_light() +
  theme(plot.caption = element_text(hjust = 0, size = 10),
        title = element_text(size = 11)) +
  labs(x = "Lobster size (mm)", 
         y = "Density", 
         fill = "Site Name",
       title = "Probability Density for Lobster Sizes Observed at the 5 Different Sites \n in 2012 and 2018",
       caption = "Figure 2: This graph shows the density of sizes of all the lobsters in 2012 and 2018 found in the five sites.")

```



### RESULTS C
#### **COMPARING MEAN LOBSTER SIZES AT MPA VS NON-MPA SITES IN 2012 AND 2018** 




```{r}
# creating a summary table for the average sizes of lobsters, the standard deviation, and sample size for mpa sites and non mpa sites for 2012 and 2018.

lob_c_sum <- lob_tidy2 %>% 
  filter(year %in% c(2012, 2018)) %>% 
  group_by(mpa_status, year) %>% 
  summarize(mean = mean(size_mm),
            sd = sd(size_mm),
            sample_size = n())

# makiing the table look nice

lob_c_sum %>% 
  kable(col.names = c("MPA Status",
                      "Year",
                      "Mean Lobster Size (mm)",
                      "Standard Deviation",
                      "Sample Size"),
        caption = "Table 1. Mean lobster size, standard deviation and sample size for MPA and non-MPA sites in 2012 and 2018.") %>% 
  kable_styling(bootstrap_options = "striped", 
                full_width = F,
                position = "left"
                ) %>%
  add_header_above(c("LOBSTER SIZE METRICS" = 5))
```

 
```{r preparing data for t tests}
# creating a data frame for lobster MPA sites in 2012
lob_mpa_2012 <- lob_tidy2 %>% 
  filter(year == "2012") %>% 
  filter(mpa_status == "MPA") %>% 
  pull(size_mm)
  
# creating a data frame for lobster MPA sites in 2018 
lob_mpa_2018 <- lob_tidy2 %>% 
  filter(year == "2018") %>% 
  filter(mpa_status == "MPA") %>% 
  pull(size_mm)

# creating a data frame for non MPA sites in 2012
lob_non_mpa_2012 <- lob_tidy2 %>% 
  filter(year == "2012") %>% 
  filter(mpa_status == "non-MPA") %>% 
  pull(size_mm)

# creating a data frame for non MPA sites in 2018
lob_non_mpa_2018 <- lob_tidy2 %>% 
  filter(year == "2018") %>% 
  filter(mpa_status == "non-MPA") %>% 
  pull(size_mm)



```


```{r}
# running four 2-sample t tests on lobster size
# 1) Ttest for mpa site for 2012 and 2018
# 2) Ttest for non mpa sites for 2012 and 2018 
# 3) Ttests for mpa and non mpa sites for 2012  
# 4) Ttest for mpa and non mpa sites for 2018


mpa_12_18_ttest <- t.test(lob_mpa_2012, lob_mpa_2018) # 1) Ttest for mpa site for 2012 and 2018



non_mpa_12_18_ttest <- t.test(lob_non_mpa_2012, lob_non_mpa_2018) # 2) Ttest for non mpa sites for 2012 and 2018



lob_mpa_non_2012 <- t.test(lob_mpa_2012, lob_non_mpa_2012) # 3) Ttests for mpa and non mpa sites for 2012  




lob_mpa_non_2018 <- t.test(lob_mpa_2018, lob_non_mpa_2018) # 4) Ttest for mpa and non mpa sites for 2018

```


```{r}

# calculating cohen's d for mpa vs. non-mpa in 2012 and 2018

mpa_12_18_cohen <- effsize::cohen.d(lob_mpa_2012, lob_mpa_2018) # comparing mpa sites b/n 2012 and 2018

nonmpa_12_18_cohen <- effsize::cohen.d(lob_non_mpa_2012, lob_non_mpa_2018) # comparing non-mpa sites b/n 2012 and 2018

mpa_non_12 <- effsize::cohen.d(lob_mpa_2012, lob_non_mpa_2012) # comparing mpa sites and non-mpa sites in 2012

mpa_non_18 <- effsize::cohen.d(lob_mpa_2018, lob_non_mpa_2018) # comparing mpa sites and non-mpa sites in 2018

```


#### 1) Mean lobster size at MPA sites in 2012 and 2018

```{r include = FALSE}
mpa_12_18_ttest
mpa_12_18_cohen
```
- Mean lobster sizes differ significantly at MPA sites between 2012 and 2018 (t(`r round(mpa_12_18_ttest$parameter, 2)`) = `r round(mpa_12_18_ttest$statistic, 2)`, p = `r round(mpa_12_18_ttest$p.value, 6)`). 
- The 95% confidence interval is (`r round(mpa_12_18_ttest$conf.int, 2)`). 
- The effect size is large (Cohen's d = `r round(mpa_12_18_cohen$estimate, 3)`). 
- Mean lobster sizes at MPA sites in 2012 and 2018 are `r round(mpa_12_18_ttest$estimate, 2)` mm, respectively. The difference in means is `r round(diff(head(mpa_12_18_ttest$estimate, 1) - head(mpa_12_18_ttest$estimate, 2)),2)` mm.



#### 2) Mean lobster size at non-MPA sites in 2012 and 2018
```{r include = FALSE}
non_mpa_12_18_ttest
nonmpa_12_18_cohen
```
- Mean lobster sizes at non-MPA sites do not differ significantly between 2012 and 2018 (t(`r round(non_mpa_12_18_ttest$parameter, 2)`) = `r round(non_mpa_12_18_ttest$statistic, 2)`, p = `r round(non_mpa_12_18_ttest$p.value, 4)`). 
- The 95% confidence interval is (`r round(non_mpa_12_18_ttest$conf.int, 2)`). 
- The effect size is negligible (Cohen's d = `r round(nonmpa_12_18_cohen$estimate, 3)`). 
- Mean lobster sizes at non-MPA sites in 2012 and 2018 are `r round(non_mpa_12_18_ttest$estimate, 2)` mm, respectively. The difference in means is `r round(diff(head(non_mpa_12_18_ttest$estimate, 1) - head(non_mpa_12_18_ttest$estimate, 2)),2)` mm.



#### 3) Mean lobster size at MPA and non-MPA sites in 2012

```{r include = FALSE}
lob_mpa_non_2012 
mpa_non_12
```
- Mean lobster sizes differ significantly between MPA and non-MPA sites in 2012 (t(`r round(lob_mpa_non_2012$parameter, 2)`) = `r round(lob_mpa_non_2012$statistic, 2)`, p = `r round(lob_mpa_non_2012$p.value, 4)`). 
- The 95% confidence interval is (`r round(lob_mpa_non_2012$conf.int, 2)`). 
- The effect size is medium (Cohen's d = `r round(mpa_non_12$estimate, 3)`). 
- In 2012, mean lobster sizes at MPA and non-MPA sites are `r round(lob_mpa_non_2012$estimate, 2)` mm, respectively. The difference in means is `r round(diff(head(lob_mpa_non_2012$estimate, 1) - head(lob_mpa_non_2012$estimate, 2)),2)` mm.

#### 4) Mean lobster size at MPA and non-MPA sites in 2018

```{r include=FALSE}
lob_mpa_non_2018
mpa_non_18
```
- Mean lobster sizes differ significantly between MPA and non-MPA sites in 2012 (t(`r round(lob_mpa_non_2018$parameter, 2)`) = `r round(lob_mpa_non_2018$statistic, 2)`, p = `r round(lob_mpa_non_2018$p.value, 14)`). 
- The 95% confidence interval is (`r round(lob_mpa_non_2018$conf.int, 2)`). 
- The effect size is small (Cohen's d = `r round(mpa_non_18$estimate, 3)`). 
- In 2018, mean lobster sizes at MPA and non-MPA sites are `r round(lob_mpa_non_2018$estimate, 2)` mm, respectively. The difference in means is `r round(diff(head(lob_mpa_non_2018$estimate, 1) - head(lob_mpa_non_2018$estimate, 2)),2)` mm.


## SUMMARY
- Isla Vista has the most increase in lobster abundance from 2012 to 2018 wth a large spike after 2016.
- There was a large decrease in lobster abundance in Carpenteria from 2017 to 2018.
- The sites with the biggest lobster size shifts from 2012 to 2018 are Isla Vista, Naples, and Carpenteria.
- Mean lobster size at MPA sites from 2012 to 2018 are significantly different.
- Mean lobster size at non MPA sites from 2012 to 2018 are not significantly different.
- Mean lobster size at MPA sites and non MPA sites in 2012 are significantly different.
- Mean lobster size at MPA sites and non MPA sites in 2018 are significantly different.

## REFERENCE


- Hadley Wickham (2017). tidyverse: Easily Install and Load the 'Tidyverse'. R
  package version 1.2.1. https://CRAN.R-project.org/package=tidyverse
- Hao Zhu (2019). kableExtra: Construct Complex Table with 'kable' and Pipe
  Syntax. R package version 1.1.0. https://CRAN.R-project.org/package=kableExtra  
- Reed D. 2019. SBC LTER: Reef: Abundance, size and fishing effort for California Spiny Lobster ( Panulirus interruptus ), ongoing since 2012. Environmental Data Initiative. https://doi.org/10.6073/pasta/a593a675d644fdefb736750b291579a0 . Dataset accessed 11/12/2019.
- Sam Firke (2019). janitor: Simple Tools for Examining and Cleaning Dirty Data. R
  package version 1.2.0. https://CRAN.R-project.org/package=janitor
- Torchiano M (2019). _effsize: Efficient Effect Size Computation_. doi:
10.5281/zenodo.1480624 (URL: https://doi.org/10.5281/zenodo.1480624), R package
version 0.7.6, <URL: https://CRAN.R-project.org/package=effsize>. 