---
title: "ESM 206 Assignment 4 Report"
author: "Ruiheng Jin and Qusai Bhaijeewala"
date: "11/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

```{r}
library(tidyverse)
library(janitor)
library(DT)
lob_abun <- read_csv("lobster_abundance_sbc_lter.csv",
                     na = "-99999") %>% 
  clean_names()
```
## INTRODUCTION


## DATA AND METHODS


## RESULTS {.tabset .tabset-fade}

### RESULTS A
#### CHANGES IN ANNUAL LOBSTER ABUNDANCE AT FIVE SITES

```{r}
# data structure: each site has several transects, each transect has ABCD replicates

# uncount lobsters so each lobster has its own row
lob_tidy <- lob_abun %>% 
  tidyr::uncount(count) %>% 
  mutate(mpa = site %in% c("IVEE", "NAPL"))

# making a new column changing TRUE or FALSE to MPA and non-MPA
lob_tidy$mpa_status <- factor(lob_tidy$mpa, labels = c("non-MPA", "MPA"))

lob_tidy$site_name <- factor(lob_tidy$site, labels = c("Arroyo Quemado", "Carpinteria", "Isla Vista", "Mohawk", "Naples"))
```

```{r}
# count total lobster count by year and site
lob_yr_site <- lob_tidy %>%
  count(year, site) %>% 
  mutate(mpa = site %in% c("IVEE", "NAPL"),
         mpa_status = factor(mpa, labels = c("non-MPA", "MPA")),
         site_name = factor(site, labels = c("Arroyo Quemado", "Carpinteria", "Isla Vista", "Mohawk", "Naples")))  

ggplot(lob_yr_site, aes(x = year, y = n)) +
  facet_wrap(site_name~mpa_status) +
  geom_line(color = "#b62b2b", size = 1) +
  geom_point(color = "#4b1313", size = 2) +
  labs(x = "", 
       y = "Lobster abundance", 
       title = "Lobster Abundance at Five Sites (2012-2018)",
       caption = "Figure 1. Lobster abundance from 5 study sites, from year 2012 to 2018.") +
  scale_x_continuous(expand = expand_scale(add = c(1, 1)),
                     breaks = c(2012, 2014, 2016, 2018)) +
  scale_y_continuous(expand = expand_scale(add = c(0, 0)),
                     limits = c(0, 1000)) +
  theme_light() +
  theme(plot.caption = element_text(hjust = 0, size = 10),
        title = element_text(size = 13)) 
# NEED TO FIGURE OUT HOW TO INCREASE GAP BETWEEN  
  
  
# geom_col(position = 'dodge') 
```


### RESULTS B
#### SHIFTS IN LOBSTER SIZE DISTRIBUTION FROM 2012 TO 2018

```{r}
# data structure: each site has several transects, and within each transect they have either A, B, C, or D replicates. The count column can has multiple counts of the same transect number and replicate, we want to make each of those counts a seperate row.

# uncount lobsters so each lobster has its own row
lob_tidy2 <- lob_abun %>% 
  tidyr::uncount(count) %>% 
  mutate(mpa = site %in% c("IVEE", "NAPL"))

# making a new column changing TRUE or FALSE to MPA and non-MPA
lob_tidy2$mpa_status <- factor(lob_tidy2$mpa, labels = c("non-MPA", "MPA"))
```

Marine Protected Areas (MPA) 
Sampling began in 2012 and is ongoing. Naples and Isla Vista

two within MPAs (Naples and Isla Vista) and three outside (Arroyo Quemado, Mohawk and Carpinteria).

```{r}
## trying mike's code out to change the axis's
lob_tidy2$site_name <- factor(lob_tidy2$site, labels = c("Arroyo Quemado", "Carpinteria", "Isla Vista", "Mohawk", "Naples"))

# filtering years of 2012 and 2018 only
lob_size <- lob_tidy2 %>% 
  filter(year %in% c(2012,2018))

#graphing size distributions of lobster size at five sites from two years, 2012, 2018
# I have to spell the sites out and maybe get rid of the legend. I wrote something in the caption as a place holder, will make it better later.

# I also want to space out the y axis a little.

ggplot(lob_size, aes(x = size_mm)) +
  geom_histogram(aes(fill = site_name)) +
  facet_grid(site ~ year, scales = "free_y") +
  theme_light() +
  theme(plot.caption = element_text(hjust = 0, size = 10),
        title = element_text(size = 13)) +
  labs(x = "Lobster size (mm)", 
         y = "Lobster Count", 
         fill = "Site Name",
       title = "Lobster size distributions observed at the 5 different sites \n in 2012 and 2018",
       caption = "Figure 2: This graph shows the size distributions of all the lobsters in 2012 and 2018 found in the five sites.")
 # theme_update(plot.title = element_text(hjust = 0.5)) +
  #geom_hline(yintercept = 6377, color = "blue")
 # annotate(geom = "text",
           #x = 1940, y = 6380,
          # label = "Land Bridges Emerge",
          # color = "black") +
 # geom_point(x = 1941, y = 6417, color = "red", size = 2) +
 # annotate("segment", x = 1960, y = 6426, xend = 1941, yend = 6417,
#           size = 0.5, arrow = arrow(), color = "red") +
#  annotate(geom = "text",
          # x = 1970, y = 6429,
          # label = "Mono Basin Diversions Begin",
         #  color = "black", size = 4) +
  #geom_ribbon(aes(ymin = 6370, ymax = lake_level_feet_above_sea_level,
  ##                color = "Light Blue", alpha = .5),
  #            show.legend = FALSE)

```

There are two marine protected area (MPA), Naples and Isla Vista. These two sites have increased their lobster counts and the sizes of the lobsters as well. The sites of Mohalk and Carpenteria have also icreased in abundance and size. The site called Arroyo Quemado has the leased increase in spiny lobster size.


### RESULTS C
#### COMPARING MEAN LOBSTER SIZES AT MPA VS NON-MPA SITES IN 2012 AND 2018 


```{r}
lob_av_size <- lob_tidy2 %>% 
  group_by(year, site) %>% 
  mutate(av_size = mean(size_mm, na.rm = TRUE)) %>% 
  filter(year == "2012")


ggplot(lob_av_size, aes(x = year, y = av_size)) +
  geom_col() +
  facet_wrap(~site) +
  theme_light() +
  theme(plot.caption = element_text(hjust = 0, size = 10),
        title = element_text(size = 13)) 
  #facet_grid(site ~ year, scales = "free_y") +
 # theme_dark() 

```

##### next graph

```{r}
lob_av_size2 <- lob_tidy2 %>% 
  group_by(year, site) %>% 
  mutate(av_size = mean(size_mm, na.rm = TRUE)) %>% 
  filter(year %in% c("2012","2018"))


ggplot(lob_av_size2, aes(x = year, y = av_size, fill = year)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(site_name ~ mpa_status) +
  scale_x_continuous(breaks = c(2012, 2018),
                     labels = c("2012", "2018")) +
  theme_light() +
  theme(plot.caption = element_text(hjust = 0, size = 10),
        title = element_text(size = 13)) +
  labs(x = "Year", 
         y = "Average Lobster Size", 
       title = "Average Lobster size observed at the 5 different sites \n in 2012 and 2018",
       caption = "Figure 3: This graph shows the Average lobster size of all the lobsters in 2012 and 2018 found in the five sites.")
```


##### Messing around with the data 

```{r}

ggplot(lob_tidy2, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_grid(site ~ year, scales = "free_y") +
  theme_light() +
  theme(plot.caption = element_text(hjust = 0, size = 10),
        title = element_text(size = 13))  
   # labs(x = "site", y = "Lobster size (mm)", 
#       title = "Lobster size distributions observed at the 5 different sites \n in 2012 and 2018")
 # theme_update(plot.title = element_text(hjust = 0.5)) +
  #geom_hline(yintercept = 6377, color = "blue")
 # annotate(geom = "text",
           #x = 1940, y = 6380,
          # label = "Land Bridges Emerge",
          # color = "black") +
 # geom_point(x = 1941, y = 6417, color = "red", size = 2) +
 # annotate("segment", x = 1960, y = 6426, xend = 1941, yend = 6417,
#           size = 0.5, arrow = arrow(), color = "red") +
#  annotate(geom = "text",
          # x = 1970, y = 6429,
          # label = "Mono Basin Diversions Begin",
         #  color = "black", size = 4) +
  #geom_ribbon(aes(ymin = 6370, ymax = lake_level_feet_above_sea_level,
  ##                color = "Light Blue", alpha = .5),
  #            show.legend = FALSE)


```


##### Making a summary table for comparing MPA vs. non-MPA, 2012 and 2018
```{r}

lob_c_sum <- lob_tidy2 %>% 
  filter(year %in% c(2012, 2018)) %>% 
  group_by(mpa_status, year) %>% 
  summarize(mean = mean(size_mm),
            sd = sd(size_mm),
            sample_size = n())

datatable(lob_c_sum)

```
 ---------------------------------------------------------------
 --------------------------------------------------------------
 T: TESTS
 
 
 
```{r}
lob_mpa_2012 <- lob_tidy2 %>% 
  filter(year == "2012") %>% 
  filter(mpa_status == "MPA") %>% 
  pull(size_mm)
  
  
lob_mpa_2018 <- lob_tidy2 %>% 
  filter(year == "2018") %>% 
  filter(mpa_status == "MPA") %>% 
  pull(size_mm)

lob_non_mpa_2012 <- lob_tidy2 %>% 
  filter(year == "2012") %>% 
  filter(mpa_status == "non-MPA") %>% 
  pull(size_mm)


lob_non_mpa_2018 <- lob_tidy2 %>% 
  filter(year == "2018") %>% 
  filter(mpa_status == "non-MPA") %>% 
  pull(size_mm)




#mpa_12_18_ttest <- t.test(lob_mpa_2012, lob_mpa_2018)

#non_mpa_12_18_ttest <- t.test(lob_non_mpa_2012, lob_non_mpa_2018)



#lob_mpa_non_2012 <- t.test(lob_mpa_2012, lob_non_mpa_2012)


#lob_mpa_non_2018 <- t.test(lob_mpa_2018, lob_non_mpa_2018)




```


```{r}
#running 2 sample t test, we will run 4
# one for mpa site for 2012 and 2018, on for non mpa for 2012 and 2018, another for mpa and non mpa for 2012, and a mpa and non mpa for 2018


mpa_12_18_ttest <- t.test(lob_mpa_2012, lob_mpa_2018) #one for mpa site for 2012 and 2018,



non_mpa_12_18_ttest <- t.test(lob_non_mpa_2012, lob_non_mpa_2018) #on for non mpa for 2012 and 2018




lob_mpa_non_2012 <- t.test(lob_mpa_2012, lob_non_mpa_2012) # another for mpa and non mpa for 2012,



lob_mpa_non_2018 <- t.test(lob_mpa_2018, lob_non_mpa_2018) # mpa and non mpa for 2018






```


#### 1) TTest for MPA sites in 2012 and 2018

```{r}
mpa_12_18_ttest
```


#### 2) TTest for NON MPA sites in 2012 and 2018
```{r}
non_mpa_12_18_ttest
```

#### 3) TTest for MPA and NON MPA sites in 2012

```{r}
lob_mpa_non_2012 
```

#### 4) TTest for MPA and NON MPA sites in 2018

```{r}
lob_mpa_non_2018
```



```{r eval=FALSE, include=FALSE}
### ntrying to do T shits
lob_2012 <- lob_tidy2 %>% 
  filter(year == "2012") %>% 
  group_by(year, mpa_status) %>% 
  summarize(mean = mean(size_mm),
            sd = sd(size_mm),
            sample_size = n())

lob_2018 <- lob_tidy2 %>% 
  filter(year ==  "2018") %>% 
  group_by(year, mpa_status) %>% 
  summarize(mean = mean(size_mm),
            sd = sd(size_mm),
            sample_size = n())





```

```{r eval=FALSE, include=FALSE}
# more trying of T stuff

lob_t <- t.test(lob_2012$size_mm, lob_2018$size_mm)
```



## SUMMARY


## REFERENCE